---
title: "Case study for bank Vasilii Marushev"
editor_options:
  markdown:
    wrap: 72
output:
  html_document:
    df_print: paged
---

```{r, include=FALSE}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(agricolae)
```

```{r, include=TRUE, warning=FALSE}
broker_list <- read.csv2("C:/Users/vmaru/Downloads/broker_list[7735].csv")
hist_list <- read.csv2("C:/Users/vmaru/Downloads/historical_sample[7726].csv")
```

## 1. IQ and age

Let's see how looks our data on boxplot. On our graph we can see that
every category is allocated on different level of IQ

```{r, include=TRUE}


ggplot(hist_list, aes(y = IQ, cost_type)) +
geom_boxplot(aes(fill = cost_type))


```

Let's prove our assumption using oneway ANOVA test, using this test we
proved that at least one group is different. It is mean that belonging
to group can influence IQ.

```{r, include = T, warning=FALSE}

model_IQ <- aov(IQ ~ cost_type, data = hist_list)

summary(model_IQ)

```

Let's determine if there's a significant difference in the average
values of different groups using Fisher's Least Significant Difference
Method. According to this test, we can conclude that our output groups
vary, suggesting that IQ affects the distribution channel

```{r, include = T, warning=FALSE}

print(LSD.test(model_IQ, "cost_type"))

```

Let's see how looks our data on boxplot. On our graph we can see that
every category is allocated on different level of Age

```{r, include=TRUE}
ggplot(hist_list, aes(y = Age, cost_type)) +
geom_boxplot(aes(fill = cost_type))

```

Let's prove our assumption using oneway ANOVA test, using this test we
proved that at least one group is different. It is mean that belonging
to group can influence IQ.

```{r, include = T, warning=FALSE}

model_Age <- aov(Age ~ cost_type, data = hist_list)

summary(model_Age)

```

We are examining a statistical test to determine if there is a
significant difference in the mean values of each group using Fisher's
Least Significant Difference Method. Based on this test, we observe that
the groups in our output are indeed different, indicating that IQ has an
influence on the distribution channel.

```{r, include = T}

print(LSD.test(model_Age, "cost_type"))

```

## 2. Relationship between IQ and age

Let's see our data on scatter plot. On this graph we can see obvious
linear positive correlation.

```{r, include = T}
ggplot(hist_list, aes(y = IQ, Age)) +
geom_point()
```

As confirmation of my word calculate correlation coefficient.

```{r, include = T}
cor(hist_list$Age, hist_list$IQ)
```

A correlation coefficient of 0.98 between IQ and Age indicates a very
strong positive linear relationship between the two variables. In this
case, as a person's age increases, their IQ tends to also increase. We
can meet with problem of multicollinearity in previous topic. It can
make it hard to pinpoint which factor is truly influencing the outcome
we're trying to understand. Because one factor influence on other or on
verse.

## 3. Relationship between income and default rate

Calculation of mean default rate for each income group.

```{r, include = T}

hist_list %>%
  group_by(Inc_Gr) %>%
  summarise(def_rate = mean(Default_flg))

```

In this part I am going to conduct Anova test to find out is here
significant difference at least between two groups.

```{r, include = T}

model_Inc <- aov(Default_flg ~ Inc_Gr, data = hist_list)

summary(model_Inc)

```

We are conducting a statistical test to determine if there are
significant differences in the mean values among multiple groups using
Fisher's Least Significant Difference Method. Our analysis identifies
three distinct groups: a) those in the first income group, b) those in
the second and third groups, and c) those in the fourth group. This
suggests a relationship exists between groups a, b, and c. Specifically,
individuals receiving social security appear to be the most susceptible
to default risk. Those with incomes below 15,000 pose a lesser risk
compared to social security recipients, although the risk remains
notable. Conversely, the group with salaries exceeding 15,000 presents a
significantly lower risk, given their average default rate of 0.02. From
an economic standpoint, we observe a correlation between higher income
levels and a reduced probability of default. This correlation can be
attributed to individuals with higher incomes having greater savings or
the ability to more easily accumulate the 30,000 required for payment.

```{r,include = T}

print(LSD.test(model_Inc, "Inc_Gr"))

```

## 4. Describing full portfolio

Let's find number of defaults and observation. We can see that we get
10000 of observation and 530 defaults

```{r, include = T}

N <- NROW(hist_list$Default_flg)
N
num_of_def <- sum(hist_list$Default_flg)
num_of_def

```

According on our previous calculation let's calculate defoalt rate.
Default rate is ratio between number of default and number of
observation.

```{r, include = T}
defolt <- num_of_def/N
defolt
```

Let's compute Loss Given Default (LGD), which represents the average
loss on a loan for a customer in the event of default. In my calculation
process, I created a new dataset that exclusively comprises data from
defaulting clients. Following this, I assigned losses for each type of
housing as outlined in the text, and then calculated the average loss in
the event of default. We can see that LGD equal to 0.65 it means that in
case of default banks lost 65% of loan in average.

```{r, include = T}
def_clients <- hist_list[hist_list$Default_flg == 1, ]

def_clients$Loss[def_clients$Housing == 3] = 0.2
def_clients$Loss[def_clients$Housing != 3] = 0.7

LGD <- mean(def_clients$Loss)
LGD


```

Let's calculate average IQ.

```{r, include = T}
mean(hist_list$IQ)

```

And average age of customers.

```{r, include = T}
mean(hist_list$Age)

```

In that part we are going to calculate most frequency type of
applicants' accommodation. The most frequency applicant living with
parents.

```{r, include = T}
freq_table_house <- hist_list %>%
  group_by(Housing)%>%
  count()

freq_table_house[freq_table_house$n == max(freq_table_house$n),]
  

```

In that part we are going to calculate most frequency type of
applicants' income. The most frequency group of clients have a salary
higher than 15 000.

```{r, include = T}

freq_table_inc <- hist_list %>%
  group_by(Inc_Gr)%>%
  count()

freq_table_inc[freq_table_inc$n == max(freq_table_inc$n),]
  
```

41.8% of our applicants have university degree

```{r, include = T}

freq_table_ed <- hist_list %>%
  group_by(Educ)%>%
  count()

freq_table_ed[freq_table_ed$n == max(freq_table_ed$n),]
  
```

## 5. Profitability

### 5.1 Calculation of profits portfolio for each distributional channel

So lets think that profit is difference between revenue and costs.
Revenue in ours case of bank is interest rate multiply loan. So we have
information, that loan is 30 000 for each debt that means we should just
multiply int. rate on 30000 in case when our client don't have a
default. Otherwise we have a costs on our product. That costs contain
cost of default it is average LGD multiply 30 000 and number of default.
But we shouldn't forget about cost of providing services for
distributional channel.

Let's calculate revenue for each distributional channel

```{r, include = T}

revenue <- hist_list %>%
  group_by(cost_type)%>% 
  filter(Default_flg == 0) %>%
  summarise(rev = sum(rate)*30000)

revenue

```

Let's calculate costs of default for each distributional channel

```{r, include = T}

cost_def <- def_clients %>%
  group_by(cost_type) %>%
  summarise(cost_on_def = sum(Loss)*30000)

cost_def

```

Let's calculate costs of providing services for each distributional
channel

```{r, include = T, warning= F}

cost_dist <- hist_list %>%
  group_by(cost_type) %>%
  summarise(n = n())

cost_dist$cost_for_loan[cost_dist$cost_type == "Broker"] = 1000
cost_dist$cost_for_loan[cost_dist$cost_type == "www"] = 200
cost_dist$cost_for_loan[cost_dist$cost_type == "Facebook"] = 300
cost_dist$total <- cost_dist$n*cost_dist$cost_for_loan

cost_dist

```

Let's calculate profit for each distributional channel.

```{r, include = T}

profit_cz <- revenue$rev - cost_def$cost_on_def - cost_dist$total

profit <- data.frame(distributional_channel = c("Broker", "Facebook", "www"),profit_cz)
profit

```

Based on our data, we can see that the most profitable distributional
channel is Facebook. I believe this channel is the most profitable
because it incurs significantly lower costs than others, and due to the
relatively small number of applicants, it is more profitable in our
specific case. However, it's essential to note that all of our channels
have some level of profitability issues. In the case of the website
(www), there is a high default rate, while in the case of the Broker
channel, the cost of every loss provided is high. Facebook seems to
strike a balance between these extremes.

## 6. Top and least profitable customers

```{r, include = F}

loan = 30000

hist_list$cost_for_loan[hist_list$cost_type == "Broker"] = 1000
hist_list$cost_for_loan[hist_list$cost_type == "www"] = 200
hist_list$cost_for_loan[hist_list$cost_type == "Facebook"] = 300

hist_list$Loss[(hist_list$Default_flg ==1)& (hist_list$Housing == 3)] = 0.2*loan
hist_list$Loss[(hist_list$Default_flg == 1) & (hist_list$Housing != 3)] = 0.7*loan
hist_list$Loss[hist_list$Default_flg == 0] = 0

hist_list$rev = hist_list$rate*loan

hist_list$rev[hist_list$Default_flg == 1] = 0

hist_list$profit = hist_list$rev - hist_list$Loss - hist_list$cost_for_loan

hist_list %>%
  filter(Housing == 3 )

```

To finding out the most profitable clients I going to use 95 quantile of
profit. It means that by this action we will find the 5% of the most
profitable clients.

```{r, include = T}

q95 <- quantile(hist_list$profit, probs = 0.95, type = 2)

top <- hist_list[hist_list$profit > q95,] 

head(top)



```

Do the same thing to finding the worst profitable applicants.

```{r, include = T}

q05 <- quantile(hist_list$profit, probs = 0.05, type = 2)

least <- hist_list[hist_list$profit < q05,]
head(least)

```

This frequency table focuses on education. Observing the table, we note
a substantial 23% higher percentage of applicants who have completed
only elementary school among the least profitable group. Additionally, a
notable difference is evident in the group of individuals with a Ph.D.
These observations suggest that education might indeed influence our
profitability. The significant variation in education levels between the
two profit groups supports this assumption.

```{r, include = T}
# freq table for the least profitable clients
Educ_com_w <- least %>% 
  group_by(Educ) %>%
  summarise(worst = n()/NROW(least)) 
# freq table for the most profitable clients
Educ_com_b <- top %>% 
  group_by(Educ) %>%
  summarise(best = n()/NROW(top)) 
# full join 
freq_table_Educ <- full_join(Educ_com_b, Educ_com_w, by = "Educ")

freq_table_Educ[is.na(freq_table_Educ)] = 0 # replacing Na

round(freq_table_Educ,2)

```

In this instance, we will construct a frequency table focusing on our
clients' income. This table presents the ratio within each income group.
Based on our dataset, there is a notable 22% difference between
customers earning more than \$30,000. Essentially, the best profitable
clients have a 22% higher representation of customers with incomes
exceeding \$30,000 compared to the least profitable clients.
Additionally, a difference is evident among clients receiving social
security---this ratio is 13% higher for the least profitable clients
when compared to the best profitable ones. For the second and third
income groups, the ratio of clients between the two groups remains
nearly identical. This observation suggests that income indeed
influences profitability. However, it's important to note that in the
previous section, we established that income groups can also influence
risk.

```{r, include = T}

# freq table for the least profitable clients
Inc_com_w <- least %>% 
  group_by(Inc_Gr) %>%
  summarise(worst = n()/NROW(least)) 
# freq table for the most profitable clients
Inc_com_b <- top %>% 
  group_by(Inc_Gr) %>%
  summarise(best = n()/NROW(top)) 
# full join 
freq_table_Inc_com <- full_join(Inc_com_b, Inc_com_w, by = "Inc_Gr")

freq_table_Inc_com[is.na(freq_table_Inc_com)] = 0 # replacing Na

round(freq_table_Inc_com,2)

```

In this section, we construct an additional frequency table based on the
type of applicants' accommodation. Initially, I anticipated a
significant difference in profitability between the best and worst
candidates living in their own apartments. However, the table reveals a
different outcome. Personally, I find this category less relevant for
our analysis, as I believe all categories, except the third one, exhibit
similar patterns. The ratio in the third category, which is 2% higher
for the worst profitable, holds negligible significance in our context.
I maintain the belief that individuals with apartments are more
reliable, as a default results in lesser loss for the bank.

```{r, include=TRUE}

# freq table for the least profitable clients
Inc_house_w <- least %>% 
  group_by(Housing) %>%
  summarise(worst = n()/NROW(least)) 
# freq table for the most profitable clients
Inc_house_b <- top %>% 
  group_by(Housing) %>%
  summarise(best = n()/NROW(top)) 
# full join 
freq_table_house_com <- full_join(Inc_house_b, Inc_house_w, by = "Housing")

freq_table_house_com[is.na(freq_table_house_com)] = 0 # replacing Na

round(freq_table_house_com,2)


```

Let's generate a frequency table that illustrates the distribution of
marriage statuses among the most and least profitable clients. Upon
reviewing our table, it becomes evident that our data align closely, and
I am inclined to believe that marital status has minimal influence on
our profitability. This conclusion is drawn from the nearly identical
ratios observed among both the best and worst profitable clients.

```{r, include=TRUE}

# freq table for the least profitable clients
Mar_com_w <- least %>% 
  group_by(MarStat) %>%
  summarise(worst = n()/NROW(least)) 
# freq table for the most profitable clients
Mar_com_b <- top %>% 
  group_by(MarStat) %>%
  summarise(best = n()/NROW(top)) 
# full join 
freq_table_Mar_com <- full_join(Mar_com_b, Mar_com_w, by = "MarStat")

freq_table_Mar_com[is.na(freq_table_Mar_com)] = 0 # replacing Na

round(freq_table_Mar_com,2)

```

# 7. Summarize

So, in this chapter, we'll summarize our findings. Firstly, I'd like to
highlight that this business may not be inherently profitable due to an
average interest rate of 0.05 and a default rate of 0.05. In the event
of default, the bank incurs an average loss of 65% of the loan amount,
totaling \$19,500 per client.

Additionally, the bank imposes substantial fees on distribution. For
example, with Broker distribution, our average profit (without
considering defaults) amounts to 500 CZK. To maintain a neutral default
rate in this channel, it should not exceed 0.025. While other
distribution channels charge lower fees, they pose a higher default risk
compared to the Broker channel. Nevertheless, the average profit,
excluding provision fees, stands at 1250 CZK.

The majority of our clients originate from the Broker channel, which
boasts a better default rate than the others. My recommendation to
enhance profitability is to negotiate with the Broker distribution
channel to reduce the broker provision, given its relatively high fee.
Furthermore, I suggest negotiating with each distributional channel to
implement a payment charge only after repayment.

If this were my bank, I would revise the scoring system and exercise
caution in providing debt. I've observed that a significant portion of
our clients struggle to afford this debt, necessitating a proactive
adjustment of our system before proceeding further."
